---
title: "CitiBike"
author: "Derek"
date: "2/17/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#4 How does weather impact ridership?
####Reading the data
```{r}
library(geosphere)
#df_names=c('SAMPLE.csv','SAMPLE2.csv')
df_names=c("201601-citibike-tripdata.csv")
ride=data.frame()
for (i in df_names){
  ride=rbind(ride,read.csv(i))
}
weather=read.csv('weather_data.csv')
#weather impact on people of different group
#snow tolerance for different ages
#temperature impact on ridership
#run it as a logistic regression
```
####Extract features
```{r}
age_group=function (age){
  if (age<=22){return ('Young Students')}
  else if (age<=30){return ('Young Professionals')}
  }

ride$distance <- distHaversine(ride[ ,c("start.station.longitude", "start.station.latitude")],ride[ ,c("end.station.longitude", "end.station.latitude")])
ride$gender=factor(ride$gender,levels=c(0,1,2),labels=c("unkown","male","female"))
ride$date=factor(format(strptime(ride$starttime,format = "%m/%d/%Y %H:%M"),"%m/%d/%Y"))
ride$age=2016-ride$birth.year

ride$birth.year=NULL
ride$starttime=NULL
ride$stoptime=NULL
ride$start.station.latitude=NULL
ride$end.station.latitude=NULL
ride$start.station.longitude=NULL
ride$end.station.longitude=NULL
ride$start.station.id=NULL
ride$end.station.id=NULL
ride$bikeid=NULL
ride$DATE=ride$date
ride$date=NULL

weather$STATION=NULL
weather$STATION_NAME=NULL
weather$DATE=factor(format(strptime(weather$DATE,format = "%Y%m%d"),"%m/%d/%Y"))

na.omit(weather)
na.omit(ride)

```
####Merge the bike dataframe with the weather dataframe, so each row of the ride information will have weather infomation for that date
```{r}
combined_df=merge(x = ride, y = weather, by = "DATE", all.x = TRUE)
library(ggplot2)
library(dplyr)
```
####Does snow/rain impact each customer group, in terms of ride time and ride displacement?
#####First, let's plot the distribution and see if there is a clear pattern
```{r}
ggplot()+geom_point(aes(x=PRCP,y=tripduration,color='Rain Impact on Customer'),combined_df[combined_df$usertype=='Customer',])+geom_point(aes(x=PRCP,y=tripduration,color='Rain Impact on Subscriber'),combined_df[combined_df$usertype=='Subscriber',])

ggplot()+geom_point(aes(x=SNOW,y=tripduration,color='Snow Impact on Customer'),combined_df[combined_df$usertype=='Customer',])+geom_point(aes(x=SNOW,y=tripduration,color='Snow Impact on Subscriber'),combined_df[combined_df$usertype=='Subscriber',])
```
#####Then, let's use regression to apporximate a curve that represents the relationship between riding activity level and weather infomation. Because there is a huge dataset, we will take a random example of 100,000 datapoints for each customer group.
```{r}
grouped_df_customer=combined_df[sample(nrow(combined_df[combined_df$usertype=='Customer',]),100000,replace=TRUE),]
grouped_df_subscriber=combined_df[sample(nrow(combined_df[combined_df$usertype=='Subscriber',]),100000,replace=TRUE),]

#model_customer_snow <- lm(tripduration ~ poly(SNOW,2), data=grouped_df_customer)
model_customer_snow <- lm(tripduration ~ SNOW, data=grouped_df_customer)

#cutomer_snow_predict=predict(model_customer_snow)

plot(tripduration~SNOW,grouped_df_customer)
curve(predict(model_customer_snow,newdata=data.frame(wt=x)),add=T)
```
####Does snow/rain impact each age group, in terms of ride time and ride displacement?
```{r}

```
####Does temeprature impact each customer group, in terms of ride time and ride displacement?
```{r}

```
####Does temperature impact each age group, in terms of ride time and ride displacement?
```{r}

```

